---
title: "Understanding Your Objective"
format: docx
editor: visual
---

## In the Beginning...

Though taboo to openly debate with a children's book writer, we think @Seuss1990 may have gotten it wrong. To be clear, we have no doubt that learning to code and conduct analytics will take you to many exciting places that are difficult to foresee. We also think a stimulating way to spend a casual afternoon is by exploring novel data sets to see what you discover. However, for those likely to have picked up this book, any good research or business project (read: likely to be efficiently successful) starts with a clear objective and well-defined deliverable.

This chapter is about three things related to the optimal function of any analytic task. The first involves figuring out exactly what needs to be done. Oftentimes you might hear seemingly straightforward questions in need of a data-based answer such as "Which patients do we do better with?", "Which staff have clients who show better progress?", or "What is the biggest driver of canceled appointments?" Just like operational definitions (e.g., citation), data definitions can be nuanced, varied, and change over time. Different ways in which we define different data components of a question will lead to different answers. Thus, the first critical analytic step is to pin down exactly what the analytic deliverable will be.

The second item this chapter is about is clearly specifying out who will be involved. This comes in two forms. First, who will be involved in actually pulling off the analytic project so that the deliverable can be, well, delivered. Second, who will use the answer once known and how will use it. As a question, what is the function of these data in the end-users workflow? To meet this function well, analytic projects need to clearly identify the roles of each team member in delivering the deliverable and how the deliverable will be used once delivered.

The third item this chapter is about is scoping your work to fulfill the identified function. Stories have been told of mythical humans who luxuriously meander through project ideation and framing, who can complete projects on their own schedule and at their own pace, and who can continuously tinker until everything is exactingly perfect. For the rest of us, deadlines are real. Dissertations need to be done. Projects need to be presented. And, deliverables need to be delivered. All of which need to be on time and to budget. Once an analytic deliverable is clearly outlined, people want to know when it will be delivered. Thus, the final section of this chapter reviews how to take the information gathered about deliverables and roles to determine how long it will take you to do the voodoo that you do so well.

## Specifying the Deliverable

Specifying the deliverable is an act of translation. Specifically, translating someone's question about the world into two components: (1) the data that can best answer the question, and (2) the analytic method that can best answer the question.

### Getting the Best Data.

Identifying the data that can best answer an analytic question might be a new skill for some behavior analysts. It is true that behavior analysts are trained in how to create operational definitions for behaviors of interest and intervention conditions (e.g., baseline vs. DRA). That is, given something we want to track or know about, behavior analysts are trained to define exactly what the behavior or environmental event looks like. In so doing, relatively precise data can be collected specific to the occurrence of a phenomenon (e.g., frequency, duration, latency, treatment fidelity). Note, however, this is a future-oriented task. I have something I am interested in. I develop a definition. I then collect data to answer that specific question (e.g., "What is the function of a behavior?" or "What is the client's current manding repertoire?"). This is the common approach in clinical and research settings.

Answering analytic questions in practice is often a past-oriented and future-oriented task. That is, given a question in need of a data-based answer, we have to look at what data we already have as well as what data would be ideal to answer the question perfectly. For example, if I am interested in knowing which patients our clinic tends to do better with, I would look at what data I have now around each patient's overall clinical presentation, what data I have now around measuring progress consistently across all patients, and what the ideal data might be for each of these areas. Assessing your data in this way lets you know: (1) what is possible with the data you have right now; (2) what are the limitations to our answer with the data we have right now; (3) what might it take to build those ideal data sets in terms of time, effort, and cost; and (4) whether collecting those ideal data sets will be worth it.

The two paragraphs above highlight that conducting analytics in practice may require behavior analysts to think differently about how they ask and answer questions with data. The questions likely asked of a data analyst or data scientist rarely will be things where the data can be collected in days or weeks for subsequent analysis. It can take months or years to build the ideal data set for a specific question. But, this doesn't mean you have to sit on your hands and do nothing while you wait. As noted above, people want answers and information that can help them make better decisions, even if those answers are not perfect.

The main takeaway of this section is that we rarely have access to data that perfectly matches a question in need of a data-based answer. This makes identifying the "best" data set a relative term. It certainly is unlikely to be the ideal data set (and this day may never come). But, by looking at what is available, we can identify what data most closely matches the question at hand, what questions are possible to answer with the data we have, and how we might want to change our data collection systems moving forward. In short, matching your data to a question is likely to shape exactly what deliverable(s) is (are) possible.

### Choosing the Best Method.

\[identifying how best to answer: Types of deliverables (e.g., dashboards, insight / learning, data product, decision needs to be made or action needs to be taken.\]

\[Identifying the primary end-user. Who is the audience? What function does the deliverable serve in their daily workflow? How precise does it need to be? Or is directional / order-of-magnitude sufficient.\]

\[Mocking up deliverable to ensure alignment among stakeholders.\]

\[Understanding repeat-ability and refreshment requirements (e.g., one-off requests, ongoing maintenance, iteratively improving projects\]

## Knowing Your Role

-   What is your role as a data analyst?

    -   Assume outside of academia

    -   Who else is on your team? SMEs

    -   Quality validation

    -   Setting expectations

-   Go from business requests to formulating deliverables, timeline, scope

## Chapter Checklist

-   Have some kind of checklist or tool - something tangible to reference

    -   Maybe action item list at the end of each section

```{r, include = FALSE}

x <- c(1,2,3,4)

```

```{r}
x
```

```{python}



```

Add a little bit about knowing your role, which includes being a content area expert. Also know what questions to ask and who to ask. You don't have to have all of the answers. Not a linear end point, it's a cyclical/iterative process.